Resources

### 2025-07-07 activities.qmd
after TG2 meeting
Revision after checks from TU, 250707

@@

* Topic-contributed paper session at Joint Statistical Meeting (JSM), 6 Aug 2025. Speaker for TG2: Marc Henrion. Topic of the talk: STRATOS Topic Group 2 – Selection of variables and functional forms in multivariable analysis: Results from a review of Covid-19 prognostic models.

>>>>

* Topic-contributed paper session at Joint Statistical Meeting (JSM), 6 Aug 2025. Speaker for TG2: Marc Henrion. Topic of the talk: STRATOS Topic Group 2 – Selection of variables and functional forms in multivariable analysis: Results from a review of Covid-19 prognostic models.[Find out more](https://ww2.amstat.org/meetings/jsm/2025/)

* Invited session at 70th Annual Meeting of the German Society for Medical Informatics, Biometrics and Epidemiology (Deutsche Gesellschaft für Medizinische Informatik, Biometrie und Epidemiologie e. V., GMDS) in Jena, Germany, 7-11 Sep 2025. Speaker for TG2: Gregor Buch. Topic of the talk: STRATOS Topic Group 2 – Selection of variables and functional forms in multivariable analysis: Results from a review of Covid-19 prognostic models.[Find out more (in German)](https://gmds2025.de/)

@@

* Invited talk at the [ROeS 2025](https://roes2025.at/index.html) (Conference of the Austro-Swiss Region of the International Biometric Society) in Graz, A, 14-18 Sep 2025. Topic of the session: Towards best practices in medical statistics: The need to implement innovative scientific solutions. Speaker: Theresa Ullmann. Topic of the talk: Neutral Comparisons Of Statistical Methods: Opportunities, Challenges, And Common Issues

### 2025-07-07 glossary.qmd
Revision after checks from TU, 250707

### Multivariable modelling

@@

### Descriptive modelling

Descriptive modeling refers to a type of statistical modeling used to summarize a dataset and identify patterns, relationships, and trends without making any predictions or inferences. The purpose of descriptive models is to capture the main associations between a dependent variable and the independent variables in the data set, typically using measures such as means, medians, and standard deviations of target variables conditional on the values of other variables and associational quantities such as regression coefficients or nonlinear (smoothing) response functions. Descriptive models are useful for gaining insight into the underlying structure of the data, but are not intended for predictions of future outcomes or for causal conclusions.*

@@

Variable selection is the process of identifying the most relevant independent variables for a given dependent variable in a statistical model. One commonly used method is stepwise regression, which involves sequentially adding or removing variables based on their statistical significance. Another method is regularization, which shrinks the coefficients of less important variables towards zero and can be implemented through techniques such as Lasso or Ridge regression. Information criteria, such as AIC or BIC, can also be used to compare different models and select the best subset of variables. Finally, some machine learning algorithms, such as decision trees and random forests, automatically perform variable selection by choosing the most informative variables for predicting the outcome.

>>>>

### Multivariable modeling

@@

### Descriptive modeling

Descriptive modeling refers to a type of statistical modeling used to summarize a dataset and identify patterns, relationships, and trends.  The purpose of descriptive models is to capture the main associations between a dependent variable and the independent variables in the data set, typically using measures such as means, medians, and standard deviations of target variables conditional on the values of other variables and associational quantities such as regression coefficients or nonlinear (smoothing) response functions. Descriptive models are useful for gaining insight into the underlying structure of the data, but are not mainly intended for predictions of future outcomes or for causal conclusions.

@@

Variable selection is the process of identifying the most relevant independent variables for a given dependent variable in a statistical model. One commonly used method is stepwise regression, which involves sequentially adding or removing variables based on their statistical significance. Another method is regularization, which shrinks the coefficients of less important variables towards zero and can be implemented through techniques such as Lasso or the Elastic Net. Information criteria, such as AIC or BIC, can also be used to compare different models and select the best subset of variables. Finally, some machine learning algorithms, such as decision trees and random forests, automatically perform variable selection by choosing the most informative variables for predicting the outcome.

### 2025-07-07 index.qmd
Revision after checks from TU, 250707

Multivariable models play a dominanting role in all kinds of empirical research, and they are used in addressing descriptive, predictive and explanatory research questions.

>>>>

Multivariable models play a dominant role in all kinds of empirical research, and they are used in addressing descriptive, predictive and explanatory research questions.

### 2025-07-07 resources.qmd
updated link
further small edits
typo corrected
Revision after checks from TU, 250707

---
title: "Software"
---

@@

**[Visualisation of non-linear modelling](https://biometrician.shinyapps.io/nonlinear_visualisation/): A shiny app developed by Daniela Dunkler and Georg Heinze**

@@

## R packages and other ressources provided by TG2 members

@@

Ressources:

@@

* Github website with R package (development version): https://gsk-biostatistics.github.io/tetrisplot

>>>>

---
title: "Resources"
---

@@

**[Bend your (sp)line](https://clinicalbiometrics.shinyapps.io/Bendyourspline/): A shiny app developed by TG2**

@@

## R packages and other resources provided by TG2 members

@@

Resources:

@@

* Github website with R package (development version): [https://gsk-biostatistics.github.io/tetrisplot](https://gsk-biostatistics.github.io/tetrisplot)
